An efficient audio coding scheme for quantitative and qualitative large scale acoustic monitoring using the sensor grid approach

### 2017-10-10 paper/article.tex
ml proof read up to conclusion

The coder's output bitrate as well as measurement error is computed for both data representations. This metric is composed of band analysis error and the additional loss yielded by quantization before the encoding process.

To evaluate the analysis error, we consider the Matlab \textit{ita\_toolbox} implementation \cite{itatoolbox2017} as a reference. We study both "fast" computation with 125~ms windows and "slow" 1~s analysis to assess the impact of time integration. Third-octave bands are computed using the proposed method as well as the \textit{ita\_toolbox} in the same experimental conditions, then averaged over time and compared to the toolbox's measurements on longer extracts. In all experiments the sound dataset is the same and is composed of 2~s extracts. The experiment is run for white noise as well as urban environmental recordings. As the analysis frames are of short duration, poor frequency resolution and spectral leakage are expected, particularly in the lower frequencies bands. For this reason we evaluate the effect of the use of windowing functions on each band measurement error in order to mitigate this loss of precision. To provide relevant statistics, these quantities are estimated with data from the 4500 audio recordings of the UrbanSound8k dataset \cite{salamon2014} described in the next section and 4500 \textit{iid.} noise sequences.

@@

In order to ensure that the proposed scheme allows the recognition of events using state of the art methods, we consider the results achieved by several recognition systems on the UrbanSound8k dataset\cite{salamon2014}. It features about 9 hours of urban environmental sounds recordings separated in 8732 wave files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, engine idling, gun shot, jackhammer, siren, street music), and distributed in 10 independent folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled at 44.1~kHz, normalized and a single channel is considered.

@@

We evaluate the global loss of information induced by the coder by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are found with a grid search. We optionally apply a discrete cosine transform to the critical band signal representation to obtain cepstra, known as Mel Frequency Cepstrum Coefficients or MFCC in the case of Mel spectrograms. The 25 first coefficients are conserved, and summarized along time with the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance operators. The feature vector is thus comprised of 275 values to reproduce available results and compare with the data produced by the proposed coding scheme. Models are trained for each setup using the 10-fold cross validation method provided by the authors of \cite{salamon2014}, that is, every combination of testing one fold on models trained with the other nine.

@@

It is important that the proposed scheme ensures a high level of recognition of acoustic events but, as the data is transmitted over the network and potentially stored, the level of intelligibility of potenital speech utterances present in the decoded stream should be as low as possible. Thus, intelligibility in decoded and reconstructed audio is also an important concern of this study.

@@

This test requires an estimate of the time-domain signal. To do so, a linearly-scaled spectrogram from the band-focused representations is estimated. This estimation can be achieved by multiplying the spectral representation by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is then approximated by either white noise spectrogram scaling or the Griffin \& Lim algorithm \cite{griffin1984} in the case of overlapping frames. The signal is then retrieved using concatenation or overlap-add depending on the overlap setting. When using overlap to compute third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.

@@

As can be expected when performing a short time analysis, high errors appear in low frequency bands where precision is poor. The slow analysis yields better estimations as it provides a globally better resolution and thus also reduces the impact of spectral leakage effects in this region. When considering white noise the error is low on the whole spectrum. Its mean over the testing samples is close to zero and its standard deviation increases in low frequencies, confirming the above-mentioned considerations. This effectiveness does not translate well to environmental sound analysis, most likely mainly due to the sound level disparities absent in the former case but omnipresent in the latter. Spectral leakage induces a correlation between close frequency components. In lower frequencies, large differences between adjacent frequency bins can thus cause important errors on third-octave bands computations. This issue is however not specific to Fourier transform based schemes as it is also encountered with a time domain filtering approach.

@@

The comparison of the error achieved by the proposed scheme to that of third-octave bands estimated by a reference time-filtering algorithm is now considered. It results in important similarities. For fast measurements and a rectangular window (Figure~\ref{fig:errormu}a), a student t-test ($p<0.05$) shows that the errors are not significantly different between 125~Hz and 12~kHz bands. For the first bands, the two methods yield different error distributions but of similar importance. Using a Hann window function results in lower error in low frequencies but adds a bias to the previously accurate band measurements. Again comparing the two computation methods shows a significant difference in the bands below 100~Hz only. A slow analysis results in statistically equivalent errors between the two methods apart from the very low frequencies: $p = 0.011$ with the rectangular window (Figure~\ref{fig:errormu}c), $p = 0.003$ at 20~Hz and $p = 0.021$ at 25~Hz for the Hann window and 66\% overlap (Figure~\ref{fig:errormu}d).\\

@@

We also examine the additional error caused by the encoding steps after obtaining the desired data representation. The only lossy operation applied is quantization, which is defined by an output word size $q$ in bits. However, it corresponds to the word size \textit{after} $\Delta$-compression, meaning that the data is in fact quantized on $2^{q-1}$ values. Since these measurements are already expressed in dB, the error $\varepsilon$ is:
\begin{equation*}
	\varepsilon = |x_q-x|
\end{equation*}
where $x_q$ and $x$ are the quantized and clean representations repectively. $x_q$ is given by:
\begin{equation*}
x_q(n) = \frac{\Delta_x}{2^{q-1}-1}\textrm{round}\left(\frac{(2^{q-1}-1)x(n)}{\Delta_x}\right), x\in \left[0, \Delta_x\right]
\end{equation*}
Figure~\ref{fig:error_q} shows an estimation of $\varepsilon$ as a function of $q$ for third-octave and Mel bands. In both cases, the mean and standard deviation seem to decrease by a factor of 2 as $q$ increases. To better understand this phenomenon, let us model $x$ as a uniform distribution such as $x\sim \textit{U}\{0, \Delta_x\}$. While this is not exact it matches our objective when using a logarithm to flatten given PDFs. The error $\varepsilon$ will follow a uniform distribution $U\sim \{0, \frac{\Delta_x}{2\times (2^{q-1}-1)}\}$. Therefore, the mean and standard deviation are:
\[
\begin{cases}
	\mu_\epsilon = \frac{\Delta_x}{4\times (2^{q-1}-1)}\\
	\sigma_\epsilon = \frac{1}{12}\frac{\Delta_x}{2\times (2^{q-1}-1)}
\end{cases}
\]
justifying the decaying ratio as $q$ increases. In practice, we assume that $\varepsilon = f(\Delta_x, \frac{1}{2^{q-1}-1})$ with the heterogeneity of the data's PDF inducing small variations.

@@

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.5\textwidth]{figures/class_mel_q.eps}
	\caption{Classification accuracy as a function of word size before encoding. The baseline on the right is computed without quantizing the representation.}
	\label{fig:class_mel_q}
\end{figure}

@@

To further analyze both representations advantages, the confusion matrix is a reliable tool. It aims at providing information regarding class-by-class accuracy and misclassification rates. We thus compute these metrics for the SVM classifier, with predictions accumulated over the ten folds in test configuration. The analogous natures of the two descriptors is highlighted by close one-versus-one differentiation performances, with a slightly lower accuracy for third-octave bands on average. Both representations yield best results for the \textit{Gun shot} class with 88.2\% for third-octave cepstra and 86.6\% for Mel cepstra. Their poorest accuracy is on the \textit{Air conditioning} class with 32.0\% and 38.1\% respectively. However, a noticeable difference between them is that on the log-frequency scale the bandwidth of Mel filters narrows as frequency increases, while third-octave are evenly distributed. The effect of this can be seen between classes \textit{Drilling} and \textit{Jackhammer}. These sounds involve important low-frequency information which can generally differentiate them, leading third-octave descriptors to perform better. Conversely, using Mel-based cepstra improves globally \textit{Air conditioning} recognition as most of its defining components are situated in higher frequencies.\\

@@

The perceptual test \footnote{Decoded speech utterances for various conditions can be listened to at \url{https://felixgontier.github.io/cense-coder/}} then evaluates the impact of analysis frame rate on intelligibility using the average intelligibility score (AIS) given by the subjects and the intelligibility ratio (IR) computed as the number of correctly transcribed words versus the number of words in the spoken utterance. Given the test conditions presented in Section\ref{sec:inintelligibility} and the number of subjects, the mean and standard deviation are computed on an average of 54 results per frame rate parameter: 12 subjects evaluating each analysis resolution 4.5 times on average.

@@

The proposed approach has the following benefits:
\begin{itemize}
	\item precision: the impact of the quantization step for 8 bits codewords is found to be very small (< 0.5dB) and thus:
	- does not impact the quality of the acoustics indicators.
	- enables the sound recognition with the same quality than directly from the spectrogram.
	\item efficiency: using the high precision parameters, the bitrate is about 1.4~kbps and 0.4~kbps for fast and slow modes respectively.
	\item privacy: at all operating modes, the reverted audio signal is found to be inintelligible according to a perceputal test.
\end{itemize}

>>>>

The coder's output bitrate as well as measurement error is computed for both data representations. This metric is composed of \ml{spectral ?} band analysis error and the additional loss yielded by quantization before the encoding process.

To evaluate the analysis error, we consider the Matlab \textit{ita\_toolbox} implementation \cite{itatoolbox2017} as a reference. We study both "fast" computation with 125~ms windows and "slow" 1~s analysis to assess the impact of time integration. Third-octave bands are computed using the proposed method as well as the \textit{ita\_toolbox} in the same experimental conditions, then averaged over time and compared to the toolbox's measurements on longer extracts. In all experiments the sound dataset is the same and is composed of 2~s extracts. The experiment is run for white noise as well as urban environmental recordings. As the analysis frames are of short duration, poor frequency resolution and spectral leakage are expected, particularly in the lower frequencies bands. For this reason we evaluate the effect of the use of windowing functions on each band measurement error in order to potentially mitigate this loss of precision. To provide relevant statistics, these quantities are estimated with data from the 4500 audio recordings of the UrbanSound8k dataset \cite{salamon2014} described in the next section and 4500 \textit{iid.} noise sequences.

@@

In order to ensure that the proposed scheme allows the recognition of events using state of the art methods, we consider the results achieved by several recognition systems on the UrbanSound8k dataset\cite{salamon2014}. It features about 9 hours of urban environmental sounds recordings separated in 8732 audio files ranging from 1 to 4 seconds each. The recordings are labeled in 10 classes (air conditioner, car horn, children playing, dog bark, drilling, engine idling, gun shot, jackhammer, siren, street music), and distributed in 10 independent folds. A method and baseline results are also provided for the classification task. It is used in all the following experiments with the exception of intelligibility computation for which a small dataset of high-quality voice recordings is preferred. Audio files are resampled at 44.1~kHz, normalized and only the \ml{left} channel is considered for further analysis.

@@

We evaluate the global loss of information induced by the coder by implementing the four classification models proposed in \cite{salamon2014}: a support vector machine with a radial basis function kernel, a random forest classifier with 500 trees, a decision tree and a k-nearest neighbors classifier with $k = 5$. Values for the SVM parameter $C$ and RBF kernel variance $\sigma^2$ are determined using a grid search. We optionally apply a discrete cosine transform to the critical band signal representation to obtain cepstra, known as Mel Frequency Cepstrum Coefficients or MFCC in the case of Mel spectrograms. The 25 first coefficients are kept, and summarized along time with several statistics such as the mean, variance, skewness, kurtosis, minimum, maximum, median, derivative mean and variance, second order derivative mean and variance operators. The feature vector is thus comprised of 275 values to reproduce available results and compare with the data produced by the proposed coding scheme. Models are trained for each setup using the 10-fold cross validation method provided by the authors of \cite{salamon2014}, that is, every combination of testing one fold on models trained with the other nine.

@@

It is important that the proposed scheme ensures a high level of recognition of acoustic events but, as the data is transmitted over the network and potentially stored, the level of intelligibility of  speech utterances potentially captured in the decoded stream should be as low as possible. Thus, intelligibility in decoded and reconstructed audio is also an important concern of this study.

@@

This test requires an implementation of the inverse operation of the coding process to provide the listener with an estimate of the time-domain signal. To do so, a linearly-scaled spectrogram from the band-focused representations is estimated. This estimation can be achieved by multiplying the spectral representation by the scaled transpose of the forward transformation matrix. A loss of resolution increasing with frequency is induced due to the shape of the transformation, as illustrated by Figure~\ref{fig:freq}. Signal phase is then approximated by either white noise spectrogram scaling or the Griffin \& Lim algorithm \cite{griffin1984} in the case of overlapping frames. The signal is then retrieved using concatenation or overlap-add depending on the overlap setting. When using overlap to compute third-octave bands, one can also avoid framing effects produced by rectangular windowing by convoluting the signal with another windowing function prior to inverse-STFT computation.

\ml{tu parles de plein d'approches alternatives et on ne sait pas de quoi tu te sers pour les resultats. Ne presente que ce que tu utilises !!}

@@

As can be expected when performing a short time analysis, strong errors appear in low frequency bands where precision is poor. The slow analysis yields better estimations as it provides a globally better resolution and thus also reduces the impact of spectral leakage effects in this region. When considering white noise the error is low on the whole frequency range. Its mean over the testing samples is close to zero and its standard deviation increases in low frequencies, confirming the above-mentioned considerations. This effectiveness does not translate well to environmental sound analysis, most likely mainly due to the sound level disparities absent in the former case but omnipresent in the latter. Spectral leakage induces a correlation between close frequency components. In lower frequencies, large differences between adjacent frequency bins can thus cause important errors on third-octave bands computations. This issue is however not specific to Fourier transform based schemes as it is also encountered with a time domain filtering approach.

@@

The comparison of the error achieved by the proposed scheme to that of third-octave bands estimated by a reference time-filtering algorithm is now considered. It results in important similarities. For fast measurements and a rectangular window (Figure~\ref{fig:errormu}a), a student t-test ($p<0.05$) shows that the errors are not significantly different between 125~Hz and 12~kHz bands. For the lower frequency bands, the two methods yield different error distributions but of similar importance. Using a Hann window function results in lower error in the low frequency range but adds a bias to the previously accurate band measurements. Again comparing the two computation methods shows a significant difference in the bands below 100~Hz only. A slow analysis results in statistically equivalent errors between the two methods apart from the very low frequencies: $p = 0.011$ with the rectangular window (Figure~\ref{fig:errormu}c), $p = 0.003$ at 20~Hz and $p = 0.021$ at 25~Hz for the Hann window and 66\% overlap (Figure~\ref{fig:errormu}d).\\

@@

The additional error caused by the encoding steps after obtaining the desired data representation is now studied. The only lossy operation applied is quantization, which is defined by an output word size $q$ in bits. However, it corresponds to the word size \textit{after} $\Delta$-compression, meaning that the data is in fact quantized on $2^{q-1}$ values. Since these measurements are already expressed in dB, the error $\varepsilon$ is:
\begin{equation*}
	\varepsilon = |x_q-x|
\end{equation*}
where $x_q$ and $x$ are the quantized and clean representations repectively. $x_q$ is given by:
\begin{equation*}
x_q(n) = \frac{\Delta_x}{2^{q-1}-1}\textrm{round}\left(\frac{(2^{q-1}-1)x(n)}{\Delta_x}\right), x\in \left[0, \Delta_x\right]
\end{equation*}
Figure~\ref{fig:error_q} shows an estimation of $\varepsilon$ as a function of $q$ for third-octave and Mel bands. In both cases, the mean and standard deviation seem to decrease by a factor of 2 as $q$ increases. To better understand this phenomenon, let us model $x$ as a uniform distribution such as $x\sim \textit{U}\{0, \Delta_x\}$. While this is not exact it matches our objective when using a logarithm to flatten given PDFs. The error $\varepsilon$ will follow a uniform distribution $U\sim \{0, \frac{\Delta_x}{2\times (2^{q-1}-1)}\}$. Therefore, the mean and standard deviation are:
\[
\begin{cases}
	\mu_\epsilon = \frac{\Delta_x}{4\times (2^{q-1}-1)}\\
	\sigma_\epsilon = \frac{1}{12}\frac{\Delta_x}{2\times (2^{q-1}-1)}
\end{cases}
\]
justifying the decaying ratio as $q$ increases. In practice, we assume that $\varepsilon = f(\Delta_x, \frac{1}{2^{q-1}-1})$ with the heterogeneity of the data's PDF inducing small variations.

@@

\begin{figure}[htbp]
	\centering
		\includegraphics[width=0.5\textwidth]{figures/class_mel_q.eps}
	\caption{Classification accuracy as a function of word size before encoding. The baseline plotted on the right is computed without quantizing the representation.}
	\label{fig:class_mel_q}
\end{figure}

@@

To further analyze both representations advantages, the confusion matrix is a reliable tool. It aims at providing information regarding class-by-class accuracy and misclassification rates. We thus compute these metrics for the SVM classifier, with predictions accumulated over the ten folds in test configuration. The analogous designs of the two descriptors is highlighted by close one-versus-one differentiation performances, with a slightly lower accuracy for third-octave bands on average. Both representations yield best results for the \textit{Gun shot} class with 88.2\% for third-octave cepstra and 86.6\% for Mel cepstra. Their poorest accuracy is on the \textit{Air conditioning} class with 32.0\% and 38.1\% respectively. However, a noticeable difference between them is that on the log-frequency scale the bandwidth of Mel filters narrows as frequency increases, while third-octave are evenly distributed. The effect of this can be seen between classes \textit{Drilling} and \textit{Jackhammer}. These sounds involve important low-frequency information which can generally differentiate them, leading third-octave descriptors to perform better. Conversely, using Mel-based cepstra improves globally \textit{Air conditioning} recognition as most of its defining components are situated in higher frequencies.\\

@@

The perceptual test \footnote{Decoded speech utterances for various conditions can be listened to at \ml{la page est terminée ?} \url{https://felixgontier.github.io/cense-coder/web/examples.html}} then evaluates the impact of analysis frame rate on intelligibility using the average intelligibility score (AIS) given by the subjects and the intelligibility ratio (IR) computed as the number of correctly transcribed words versus the number of words in the spoken utterance. Given the test conditions presented in Section\ref{sec:inintelligibility} and the number of subjects, the mean and standard deviation are computed on an average of 54 results per frame rate parameter: 12 subjects evaluating each analysis resolution 4.5 times on average.

@@

The proposed approach has the following benefits:
\begin{itemize}
	\item precision: the impact of the quantization step for 8 bits codewords is found to be very small (< 0.5dB) and thus:
	- does not impact the quality of the acoustics indicators.
	- enables the sound recognition with the same quality than directly from the spectrogram.
	\item efficiency: using the high precision parameters, the bitrate is about 1.4~kbps and 0.4~kbps for fast and slow modes respectively.
	\item privacy: at all operating modes, the reverted audio signal is found to be inintelligible according to a perceputal test done using clean speech recorded at very low background level. The sensors being typically hanged at 4 meters high in a usually very noisy environment, we believe that the intelligibility risk is negligible.
\end{itemize}
